---
title: "analysis_1"
output: html_document
date: "2024-01-20"
---

# Setup and load files 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load("tidyverse", "readxl", "ade4", "psych", "rstatix", "caret", "harrietr", "pROC", "sjPlot", "emmeans", "stringi")
```

```{r}
#load items df and create relevant cols
items <- read_xlsx("symptoms.xlsx", sheet = 1)
items[items == "NA"] <- NA
items$item <- str_c(items$SCALE_MEASUREMENT, " ", items$POINT_SCALE)
items$item[90:101] <- items$SCALE_MEASUREMENT[90:101]
items <- items %>% dplyr::select(!c("SCALE_MEASUREMENT", "POINT_SCALE"))

items$NAME <- str_c(items$CODE, items$NAME, sep = " - ")

#load symptom df
symp <- read_xlsx("symptoms.xlsx", sheet = 2)
symp[symp == "NA"] <- NA
symp$n <- str_c("S", symp$n)

#load nosology category df 
nos_cat <- read_xlsx("symptoms.xlsx", sheet = 3)

#load symptom df with count per instrument of compound and specific symptom rep
symp_n <- read_xlsx("symptoms.xlsx", sheet = 4)

#load df with symptoms for jaccard index calc
jac <- read_xlsx("symptoms.xlsx", sheet = 5)
jac$n[1:9] <- str_c(0, jac$n[1:9])
jac$n <- str_c("S", jac$n)
jac <- jac[,1:6]

jac_wt <- jac #with c/s indication 

jac[,3:6][jac[,3:6] == 2] <- 1
```
# Manual overlap analysis 
## 1. Instrument overlap based on symptom categeroies 
### 1.1 DSM-5 and ICD-11 coverage of each other 
```{r}
nos <- items[items$SCALE == "DSM-5" | items$SCALE =="ICD-11", 1:5]

dsm_n_items <- nrow(items[items$SCALE == "DSM-5",])
icd_n_items <- nrow(items[items$SCALE == "ICD-11",])

nos_cat_o <- data.frame(symp_cat = c(unique(nos$SUB_SCALES), "not in nosologies"), 
                          DSM_5 = c((1/dsm_n_items*100),
                                    (1/dsm_n_items*100),
                                    (1/dsm_n_items*100),
                                    (1/dsm_n_items*100),
                                    (1/dsm_n_items*100), NA), 
                          ICD_11 = c((2/icd_n_items*100),
                                     (1/icd_n_items*100),
                                     (1/icd_n_items*100),
                                     (2/icd_n_items*100),
                                     (1/icd_n_items*100), NA))
```

### 1.2 Items in instruments that cover symptom catgories from nosologies 
```{r}
panss_n_items <- nrow(items[items$SCALE == "PANSS",])
ss_n_items <- nrow(items[items$SCALE == "SANS_SAPS",])

cat_o <- pivot_wider(as.data.frame(with(nos_cat, table(SCALE, NOS_CAT))), names_from = "SCALE", values_from = Freq)

cat_o$`DSM-5` <- cat_o$`DSM-5`/dsm_n_items*100
cat_o$`ICD-11` <- cat_o$`ICD-11`/icd_n_items*100
cat_o$PANSS <- cat_o$PANSS/panss_n_items*100
cat_o$SANS_SAPS <- cat_o$SANS_SAPS/ss_n_items*100

cat_o[,2:5] <- round(cat_o[,2:5], digits = 2)

```


## 2. Symptom generation 
### 2.1 Symptom appearance in scales 
```{r}
jac$scales_n <- rowSums(jac[3:6])

symp_sum <- as.data.frame(table(jac$scales_n))
symp_sum$perc_symp <- symp_sum$Freq/nrow(jac)*100
colnames(symp_sum) <- c("n_scales", "n_symp", "perc_symp")

mean(jac$scales_n)

jac_wt$scales_n <- jac$scales_n
```

### 2.2 Distribution of idiosyncratic, specific and compound symptoms per instrument 
```{r}
ins_sum <- data.frame(
  scale = c("DSM_5", "ICD_11", "PANSS", "SANS_SAPS"), 
  n_items = c(dsm_n_items, icd_n_items, panss_n_items, ss_n_items), 
  n_symp = c(colSums(jac[3:6])), 
  perc_com = c(nrow(jac_wt[jac_wt$DSM_5 == 1,]), 
               nrow(jac_wt[jac_wt$ICD_11 == 1,]),
               nrow(jac_wt[jac_wt$PANSS == 1,]),
               nrow(jac_wt[jac_wt$SANS_SAPS == 1,])),
  perc_spec = c(nrow(jac_wt[jac_wt$DSM_5 == 2,]),
                nrow(jac_wt[jac_wt$ICD_11 == 2,]),
                nrow(jac_wt[jac_wt$PANSS == 2,]),
                nrow(jac_wt[jac_wt$SANS_SAPS == 2,])),
  perc_idio = c(nrow(jac[jac$scales_n == 1 & jac$DSM_5 >0, ]),
                nrow(jac[jac$scales_n == 1 & jac$ICD_11 >0, ]),
                nrow(jac[jac$scales_n == 1 & jac$PANSS >0, ]),
                nrow(jac[jac$scales_n == 1 & jac$SANS_SAPS >0, ])))

ins_sum$perc_com <- ins_sum$perc_com/ins_sum$n_symp*100
ins_sum$perc_spec <- ins_sum$perc_spec/ins_sum$n_symp*100
ins_sum$perc_idio <- ins_sum$perc_idio/ins_sum$n_symp*100
```



## 3. Overlap estimation
### 3.1 Jaccard index 
```{r}
# DSM_5 overlap 
a1<-1-(dist.binary(matrix(c(jac$DSM_5,jac$ICD_11),nrow=2,byrow=T),method=1)^2)
b1<-1-(dist.binary(matrix(c(jac$DSM_5,jac$PANSS),nrow=2,byrow=T),method=1)^2)
c1<-1-(dist.binary(matrix(c(jac$DSM_5,jac$SANS_SAPS),nrow=2,byrow=T),method=1)^2)
DSM_5<-c(1,a1,b1,c1)

# ICD_11 overlap 
a2<-1-(dist.binary(matrix(c(jac$ICD_11,jac$DSM_5),nrow=2,byrow=T),method=1)^2)
b2<-1-(dist.binary(matrix(c(jac$ICD_11,jac$PANSS),nrow=2,byrow=T),method=1)^2)
c2<-1-(dist.binary(matrix(c(jac$ICD_11,jac$SANS_SAPS),nrow=2,byrow=T),method=1)^2)
ICD_11<-c(a2, 1, b2, c2)

#PANSS overlap 
a3<-1-(dist.binary(matrix(c(jac$PANSS,jac$DSM_5),nrow=2,byrow=T),method=1)^2)
b3<-1-(dist.binary(matrix(c(jac$PANSS,jac$ICD_11),nrow=2,byrow=T),method=1)^2)
c3<-1-(dist.binary(matrix(c(jac$PANSS,jac$SANS_SAPS),nrow=2,byrow=T),method=1)^2)
PANSS<-c(a3, b3, 1, c3)

#SANS/SAPS overlap
a4<-1-(dist.binary(matrix(c(jac$SANS_SAPS,jac$DSM_5),nrow=2,byrow=T),method=1)^2)
b4<-1-(dist.binary(matrix(c(jac$SANS_SAPS,jac$ICD_11),nrow=2,byrow=T),method=1)^2)
c4<-1-(dist.binary(matrix(c(jac$SANS_SAPS,jac$PANSS),nrow=2,byrow=T),method=1)^2)
SANS_SAPS<-c(a4,b4,c4, 1)

#table with all similarity values
jac_mat = matrix(nrow=4, ncol=4) 
colnames(jac_mat) <- c("DSM_5",	"ICD_11",	"PANSS",	"SANS_SAPS")
rownames(jac_mat) <- c("DSM_5",	"ICD_11",	"PANSS",	"SANS_SAPS")
jac_mat[1,]<-DSM_5
jac_mat[2,]<-ICD_11
jac_mat[3,]<-PANSS
jac_mat[4,]<-SANS_SAPS
isSymmetric(jac_mat)
```


```{r}
jac_mat[jac_mat == 1] <- 0 # replace diagonal with 0
#col means 
dsm_5_m <- mean(jac_mat[2:4,1])
icd_11_m <- mean(jac_mat[c(1,3,4),2])
panss_m <- mean(jac_mat[c(1,2,4),3])
sans_saps_m <- mean(jac_mat[c(1,3,3),4])

tot_m <- mean(dsm_5_m, icd_11_m, panss_m, sans_saps_m) 

corr <- 
  data.frame(
    scale = row.names(jac_mat), 
    mean = c(dsm_5_m, icd_11_m, panss_m, sans_saps_m), 
    length1 = length1<-c(5,7,30,59), 
    length2 = 
      c(nrow(jac[jac$DSM_5 == 1, ]),
        nrow(jac[jac$ICD_11 == 1, ]),
        nrow(jac[jac$PANSS == 1, ]),
        nrow(jac[jac$SANS_SAPS == 1,])))
```

### 3.2 Visualization, including symptom type (compound or specific)
#### 3.2.1 Prep df for vizualition 
```{r}
jac_wt$n <- factor(jac_wt$n, ordered = T)
jac_viz <- pivot_longer(jac_wt, cols = c("DSM_5", "ICD_11", "PANSS", "SANS_SAPS"))
jac_viz <- jac_viz %>% filter(!c(value == 0))
jac_viz$value[jac_viz$value == "1"] <- "compound"
jac_viz$value[jac_viz$value == "2"] <- "specific"

jac_viz$name <- ordered(jac_viz$name, levels = c("DSM_5", "ICD_11", "SANS_SAPS", "PANSS"))
jac_viz$ins_num <- as.numeric(jac_viz$name)
```

#### 3.2.2 Prep df angles 
```{r}
angle_df <- data.frame(id = c(1:63), 
                   angle = NA, 
                   hjust = NA)

nS <- length(jac$n) #n symptoms 
angle_df$angle <- 90-360*angle_df$id/nS
angle_df$hjust <- ifelse(angle_df$angle < -90, 1, 0)
angle_df$angle <- ifelse(angle_df$angle< -90, angle_df$angle+180, angle_df$angle)
angle_df$n <- jac_wt$n

jac_viz <- merge(jac_viz, angle_df)
colnames(jac_viz) <- 
  c("n", "symptom", "scales_n", "scale", "type", "ins_num", "id", "angle", "hjust")
```

#### 3.2.3 Prep colours for viz
```{r}
## fill colour 
dp$fill <- str_c(dp$scale, dp$type, sep = "-")

display.brewer.all(type = "seq")
or <- brewer.pal(9, "Oranges")
bu <- brewer.pal(9, "Blues")
pi <- brewer.pal(9, "PuRd")
gr <- brewer.pal(9, "Greens")

col1 <- c(pi[5], gr[5], or[4], bu[5])
s <- c("DSM_5", "ICD_11", "SANS_SAPS", "PANSS")
col1 <- setNames(col, s)

fil1 <- c(pi[c(3,5)], gr[c(3,5)], or[c(2,4)], bu[c(3,5)])
sf <- c("DSM_5-compound", 'DSM_5-specific', "ICD_11-compound", "ICD_11-specific", 
       "SANS_SAPS-compound", "SANS_SAPS-specific", 
       "PANSS-compound", "PANSS-specific")
fil1 <- setNames(fil1, sf)
```


#### 3.2.4 Plot viz 
```{r}
p1 <- ggplot(jac_viz, 
             aes(x= factor(n), y=ins_num, group=n, colour = scale, shape = type, rev=F))+
  # Lines and background
  geom_line() + 
  xlab("") +
  ylab("") +
  geom_hline(yintercept = 1:4, colour = "grey80", linewidth = .2) +
  geom_vline(xintercept = 1:63, colour = "grey80", linewidth = .2) +
  geom_line(colour="grey60", linetype = 1) +
  geom_rect(xmin=-Inf,xmax=Inf,ymin=-Inf,ymax=0.99,fill="white", color=NA) +
  geom_rect(xmin=-Inf,xmax=Inf,ymin=4.01,ymax=6,fill="white", color=NA) +
  # Points 
  geom_point(size=3, stroke = 0.9, fill = "white") +
  scale_shape_manual(values = c(21,19)) + 
  # Theme
  coord_polar() +
  theme(axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        panel.grid = element_blank(),
        panel.background = element_blank(),
        panel.border = element_blank(),
        plot.background = element_blank(),
        plot.margin = unit(c(0,0,0,1.1), "cm"),
        legend.margin = margin(-0.5,0,0,0, unit = "cm"),
        legend.position=c(0,0.8),
        legend.text = element_text(size = 7),
        legend.key.size = unit(0.4, "cm"),
        legend.key = element_rect(fill = "white")) +
  scale_y_continuous(limits=c(-4,5), expand=c(0,0), breaks=1:4) + 
  # Legends and labels 
  guides(color = guide_legend(title = "", ncol = 1), 
         shape = guide_legend(title = "", ncol = 1),
         fill = "none", alpha = "none") + 
  #axis tick labels
  geom_text(jac_viz, 
            mapping = aes(x= factor(n, levels = angle_df$n),y= 4.5, 
                          label= n, angle = angle, hjust = hjust),size=2.4, 
            inherit.aes = F, show.legend = F) ;p1
```
#### 3.2.5 Plot symptom names 
```{r}
#shorten names of symptoms 
jac$symptoms[jac$n == "S36" ] <- "Unable to complete tasks"
jac$symptoms[jac$n == "S38" ] <- "No enjoyment from activities"
jac$symptoms[jac$n == "S39" ] <- "No interest in activities"
jac$symptoms[jac$n == "S47" ] <- "Aggressive and agitated behaviour"
jac$symptoms[jac$n == "S48" ] <- "Repetetive, stereotyped behaviour"
jac$symptoms[jac$n == "S57" ] <- "Avoids interaction during interview"


jac$posx <-  rep(c(-0.9, 2.8, 6.25, 10.15, 13.8, 18.2, 22.35), each = 9)
jac$posy <- rep(9:1, 7)

p2 <- ggplot(jac, aes(x = posx, y = posy)) + 
  geom_text(mapping = 
              aes(label = str_c(n, symptoms, sep = " "), 
                  hjust = "left"), size = 1.4)+
  expand_limits(x = c(1,25), y = c(1,45))+
  theme(panel.background =element_blank(), 
        axis.text = element_blank(), 
        axis.ticks = element_blank(),
        axis.title = element_blank(), 
        plot.margin = unit(c(0,0,0,-0.9), "cm"));p2
```

#### 3.2.6 Save plot ad pdf 
```{r}
ggsave("fried_plot2.pdf", plot = p1)
ggsave("fried_text1.pdf", plot = p2)
ggsave("fried_c1.pdf", plot = p3)

```

## 4. Sentence Transformer prep 
### 4.1 Create dataframe with item pairs and how many symptoms they share 
```{r}
st <- pivot_longer(symp, cols = c(3:15), names_to = c("scale", "del", "type"), names_sep = "_*_", values_to = "code", values_drop_na = T)
st <- st[c(1:3, 5:6)]

st_mat <- pivot_wider(st[c(1,5,4)], names_from = c("code"), values_from = ("type"))
                      
st_mat <- as.data.frame(st_mat)
rownames(st_mat) <- st_mat$n
st_mat <- st_mat[2:102]

st_mat[st_mat == "compound"] <- 1
st_mat[st_mat == "specific"] <- 2

st_count <- pairwiseCount(st_mat)
```

### 4.2 Binary classification evaluator dataframe
```{r}
#version1: every pair once
BC_eval_df <- melt_dist(st_count)

colnames(BC_eval_df) <- c("item1", "item2", "count")
# 1 is awarded to each pair of items from same sub-scale in sans/saps 
for(row in 1:nrow(BC_eval_df)){
  if ((items$SUB_SCALES[items$CODE == (BC_eval_df$item1[row])] == 
      items$SUB_SCALES[items$CODE == (BC_eval_df$item2[row])]) & 
      items$SCALE[items$CODE == (BC_eval_df$item2[row])] == "SANS_SAPS" &
      items$SCALE[items$CODE == (BC_eval_df$item1[row])] == "SANS_SAPS"){
    BC_eval_df$count[row] <- BC_eval_df$count[row] + 1
  }
}

BC_eval_df$bc_overlap[BC_eval_df$count == 0] <- 0
BC_eval_df$bc_overlap[BC_eval_df$count > 0] <- 1
```

```{r}
#version2: every pair twice
BC_eval_df2 <- cor_gather(st_count)

colnames(BC_eval_df2) <- c("item1", "item2", "count")
# 1 is awarded to each pair of items from same sub-scale in sans/saps 
for(row in 1:nrow(BC_eval_df2)){
  if ((items$SUB_SCALES[items$CODE == (BC_eval_df2$item1[row])] == 
      items$SUB_SCALES[items$CODE == (BC_eval_df2$item2[row])]) & 
      items$SCALE[items$CODE == (BC_eval_df2$item2[row])] == "SANS_SAPS" &
      items$SCALE[items$CODE == (BC_eval_df2$item1[row])] == "SANS_SAPS"){
    BC_eval_df2$count[row] <- BC_eval_df2$count[row] + 1
  }
}

BC_eval_df2$bc_overlap[BC_eval_df2$count == 0] <- 0
BC_eval_df2$bc_overlap[BC_eval_df2$count > 0] <- 1
```

```{r}
length(BC_eval_df$bc_overlap[BC_eval_df$bc_overlap == 1]) #493
length(BC_eval_df$bc_overlap[BC_eval_df$bc_overlap == 0]) #4557
```


### 4.3 Export files
```{r}
write_csv(BC_eval_df, "bc_eval_df.csv")

writeLines(items$item, "items.txt")
writeLines(items$CODE, "code.txt")
writeLines(items$SCALE, "scale.txt")
writeLines(items$NAME), "name.txt")
```


# NLP overlap analysis 

## 1. Load cosine scores from pre-trained models
```{r}
cossco1 <- read.csv("cossco1.csv", row.names = 1)
cossco2 <- read.csv("cossco2.csv", row.names = 1)
cossco3 <- read.csv("cossco3.csv", row.names = 1)
```

## 2. df with cosine scores between all pairs for the three models 
```{r}
# Long dataframe for model 1, 2 and 3
cos1u <- as.data.frame(melt_dist(cossco1))
colnames(cos1u) <- c("item1", "item2", "cos_m1")
cos2u <- as.data.frame(melt_dist(cossco2))
colnames(cos2u) <- c("item1", "item2", "cos_m2")
cos3u <- as.data.frame(melt_dist(cossco3))
colnames(cos3u) <- c("item1", "item2", "cos_m3")

# Merge dfs so each models calculated cosine scores gets a column 
cosu <- merge(cos1u, cos2u, sort = F)
cosu <- merge(cosu, cos3u, sort = F)

# append instrument for each item 
cosu <- merge(cosu, items[c(1,3)], by.x = "item1", by.y = "CODE")
colnames(cosu)[6] <- 'item1_scale'
cosu <- merge(cosu, items[c(1,3)], by.x = "item2", by.y = "CODE")
colnames(cosu)[7] <- 'item2_scale'

# column indicating if the pairs is from the same instrument
cosu$WA <- ifelse(cosu$item1_scale == cosu$item2_scale, "within", "across")

# append manual overlap 
cosu <- merge(cosu, BC_eval_df2[c(1,2,4)], by.x = c("item1", "item2"), by.y = c("item1", "item2"))

cosu <- cosu %>% 
  mutate_at(c("item1", "item2", "item1_scale", "item2_scale", "WA", "bc_overlap"), as.factor)
cosu$bc <- ordered(cosu$bc_overlap)

cosu <- cosu[-9]
```

## 3. Implementing threshholds and calculating evaluation metrics for the models
### 3.1 df for threhold and metrics 
```{r}
metrics <- data.frame(name = c("cos_m1", "cos_m2", "cos_m3"), 
                      G_means_thr = NA, 
                      Sensitivity = NA, 
                      Specificity = NA, 
                      G_mean = NA, 
                      Balanced_Accuracy = NA, 
                      Precision = NA, 
                      F1 = NA)
```

### 3.2 g-means threhold and metrics 
```{r}
#model 1
m1_roc <- roc(cosu$bc, cosu$cos_m1)
m1_roc_df <- coords(m1_roc, x = "all", ret = "all")
m1_roc_df$G_mean <- sqrt(m1_roc_df$sensitivity*m1_roc_df$specificity)

metrics$G_means_thr[1] <- m1_roc_df$threshold[m1_roc_df$G_mean==max(m1_roc_df$G_mean)]
metrics$Sensitivity[1] <- m1_roc_df$sensitivity[m1_roc_df$G_mean==max(m1_roc_df$G_mean)]
metrics$Specificity[1] <- m1_roc_df$specificity[m1_roc_df$G_mean==max(m1_roc_df$G_mean)]
metrics$G_mean[1] <- m1_roc_df$G_mean[m1_roc_df$G_mean==max(m1_roc_df$G_mean)]
metrics$Balanced_Accuracy[1] <- (metrics$Sensitivity[1]+metrics$Specificity[1])*0.5
metrics$Precision[1] <- m1_roc_df$precision[m1_roc_df$G_mean==max(m1_roc_df$G_mean)]
metrics$F1[1] <- (2*metrics$Precision[1]*metrics$Sensitivity[1])/(metrics$Precision[1]+ metrics$Sensitivity[1])

#model 2
m2_roc <- roc(cosu$bc, cosu$cos_m2)
m2_roc_df <- coords(m2_roc, x = "all", ret = "all")
m2_roc_df$G_mean <- sqrt(m2_roc_df$sensitivity*m2_roc_df$specificity)

metrics$G_means_thr[2] <- m2_roc_df$threshold[m2_roc_df$G_mean==max(m2_roc_df$G_mean)]
metrics$Sensitivity[2] <- m2_roc_df$sensitivity[m2_roc_df$G_mean==max(m2_roc_df$G_mean)]
metrics$Specificity[2] <- m2_roc_df$specificity[m2_roc_df$G_mean==max(m2_roc_df$G_mean)]
metrics$G_mean[2] <- m2_roc_df$G_mean[m2_roc_df$G_mean==max(m2_roc_df$G_mean)]
metrics$Balanced_Accuracy[2] <- (metrics$Sensitivity[2]+metrics$Specificity[2])*0.5
metrics$Precision[2] <- m2_roc_df$precision[m2_roc_df$G_mean==max(m2_roc_df$G_mean)]
metrics$F1[2] <- (2*metrics$Precision[2]*metrics$Sensitivity[2])/(metrics$Precision[2]+ metrics$Sensitivity[2])

#model 3
m3_roc <- roc(cosu$bc, cosu$cos_m3)
m3_roc_df <- coords(m3_roc, x = "all", ret = "all")
m3_roc_df$G_mean <- sqrt(m3_roc_df$sensitivity*m3_roc_df$specificity)

metrics$G_means_thr[3] <- m3_roc_df$threshold[m3_roc_df$G_mean==max(m3_roc_df$G_mean)]
metrics$Sensitivity[3] <- m3_roc_df$sensitivity[m3_roc_df$G_mean==max(m3_roc_df$G_mean)]
metrics$Specificity[3] <- m3_roc_df$specificity[m3_roc_df$G_mean==max(m3_roc_df$G_mean)]
metrics$G_mean[3] <- m3_roc_df$G_mean[m3_roc_df$G_mean==max(m3_roc_df$G_mean)]
metrics$Balanced_Accuracy[3] <- (metrics$Sensitivity[3]+metrics$Specificity[3])*0.5
metrics$Precision[3] <- m3_roc_df$precision[m3_roc_df$G_mean==max(m3_roc_df$G_mean)]
metrics$F1[3] <- (2*metrics$Precision[3]*metrics$Sensitivity[3])/(metrics$Precision[3]+ metrics$Sensitivity[3])
```

### 3.3 Defining 0/1 classification for each pair based on threhold 
```{r}
cosu$m1_bc_g <- ifelse(cosu$cos_m1>metrics$G_means_thr[1], 1, 0)
cosu$m2_bc_g <- ifelse(cosu$cos_m2>metrics$G_means_thr[2], 1, 0)
cosu$m3_bc_g <- ifelse(cosu$cos_m2>metrics$G_means_thr[3], 1, 0)
```

### 3.4 Adding manual nosology category to df 
```{r}
cosu <- merge(cosu, nos_cat[2:3], by.x = "item2", by.y = "ITEMS")
cosu <- merge(cosu, nos_cat[2:3], by.x = "item1", by.y = "ITEMS")


cosu$nos_cat <- ifelse(cosu$NOS_CAT.x == cosu$NOS_CAT.y, cosu$NOS_CAT.x, 
       str_c(cosu$NOS_CAT.x, cosu$NOS_CAT.y, sep = " - "))
cosu <- cosu[-c(13:14)]

for (row in 1:nrow(cosu)){
  if (grepl(" - ", cosu$nos_cat[row])){
    cosu$nos_cat[row] <- str_c(str_sort(unlist(strsplit(cosu$nos_cat[row], 
                                                           split = " - "))), 
                                  collapse = " - ")
  }
  else (cosu$nos_cat[row] <- cosu$nos_cat[row])
}

cosu$nos_cat_WA <- ifelse(grepl(" - ", cosu$nos_cat), "across", "within")
```

## 4. Comparing the models 
### 4.1 Plotting cosine similairy distribution for manual classification and withn/across overlap
```{r}
plot(cosu$cos_m1, cosu$WA)
plot(cosu$cos_m2, cosu$WA)
plot(cosu$cos_m3, cosu$WA)

plot(cosu$cos_m1, cosu$bc)
plot(cosu$cos_m2, cosu$bc)
plot(cosu$cos_m3, cosu$bc)
```

### 4.2 Measures of spread etc of cosine scores for each model 
```{r}
cosu %>% 
  pivot_longer(cols = c(3:5)) %>% 
  ggplot()+
  geom_boxplot(aes(x = value, y = bc, alpha = 0.2)) +
  facet_wrap(~name, nrow = 3 ) +
  geom_vline(data = metrics, aes(xintercept = G_means_thr))

```


### 4.3 Density plots for each model, with threhshold line
```{r}
cosu %>% 
  pivot_longer(cols = c(3:5)) %>% 
  ggplot()+
  geom_density(aes(x = value, after_stat(count), color = bc, fill = bc, alpha = 0.2)) +
  facet_wrap(~name)+ 
  geom_vline(data = metrics, aes(xintercept = G_means_thr))
```


### 4.4 Compare within vesus across scale overlap 
```{r}
lm1 <- lm(cos_m1 ~ bc*WA, data = cosu)
lm2 <- lm(cos_m2 ~ bc*WA, data = cosu)
lm3 <- lm(cos_m3 ~ bc*WA, data = cosu)

summary(lm1)
summary(lm2)
summary(lm3)
emmeans(lm2, list(pairwise ~ WA*bc),adjust = 'bonferroni')
emmip(lm2, WA~bc, CIs = T)



lm2c <- lm(cos_m2 ~ nos_cat*m2_bc_g, data = cosu[cosu$nos_cat_WA == "within",])
summary(lm2c)
emmeans(lm2c, list(pairwise ~ nos_cat),adjust = 'bonferroni')
emmip(lm2c, nos_cat ~ m2_bc_g, CIs = T)
```

## 5. Model 2: instrument overlap 
### 5.1 df with model 2
```{r}
cosu_m2 <- cosu[c(1,2,4,11,9,6:8)]

cosu_m2 <- merge(cosu_m2, items[c(3:4)], by.x = "item2", by.y = "CODE")
cosu_m2 <- merge(cosu_m2, items[c(3:4)], by.x = "item1", by.y = "CODE")

colnames(cosu_m2)[c(1:2,9:10)] <- c("code1", "code2", "item2", "item1")

cosu_m2$item1_scale <- as.character(cosu_m2$item1_scale)
cosu_m2$item2_scale <- as.character(cosu_m2$item2_scale)

cosu_m2$ins <- ifelse((cosu_m2$item1_scale %in% c("DSM-5", "ICD-11")) &
                        (cosu_m2$item2_scale %in% c("DSM-5", "ICD-11")), "nos", 
                      ifelse(cosu_m2$item1_scale %in% c("PANSS", "SANS_SAPS") &
                        (cosu_m2$item2_scale %in% c("PANSS", "SANS_SAPS")), "scales",
                      "nos_scale"))

cosu_m2$scale_c <- str_c(cosu_m2$item1_scale, cosu_m2$item2_scale, sep = "-")
```

### 5.2 count, similar and disimilar pairs 
```{r}
length(cosu_m2$m2_bc_g[cosu_m2$m2_bc_g ==1])
length(cosu_m2$m2_bc_g[cosu_m2$m2_bc_g ==0])
length(cosu_m2$bc[cosu_m2$bc ==1])
length(cosu_m2$bc[cosu_m2$bc ==0])


length(cosu_m2$WA[cosu_m2$WA =="within"])
length(cosu_m2$WA[cosu_m2$WA =="across"])

```

```{r}
cosu_m2 <- merge(cosu_m2, nos_cat[2:3], by.x = "code2", by.y = "ITEMS")
cosu_m2 <- merge(cosu_m2, nos_cat[2:3], by.x = "code1", by.y = "ITEMS")


cosu_m2$nos_cat <- ifelse(cosu_m2$NOS_CAT.x == cosu_m2$NOS_CAT.y, cosu_m2$NOS_CAT.x, 
       str_c(cosu_m2$NOS_CAT.x, cosu_m2$NOS_CAT.y, sep = " - "))
cosu_m2 <- cosu_m2[-c(13:14)]

for (row in 1:nrow(cosu_m2)){
  if (grepl(" - ", cosu_m2$nos_cat[row])){
    cosu_m2$nos_cat[row] <- str_c(str_sort(unlist(strsplit(cosu_m2$nos_cat[row], 
                                                           split = " - "))), 
                                  collapse = " - ")
  }
  else (cosu_m2$nos_cat[row] <- cosu_m2$nos_cat[row])
}

cosu_m2$nos_cat_WA <- ifelse(grepl(" - ", cosu_m2$nos_cat), "across", "within")
```



### 5.3 instrument overlap 
#### 5.3.1 Matrix with cosine scores from model 2, NA in all cells below threhold 
```{r}
m2_mat <- as.matrix(cossco2)
m2_mat[m2_mat < metrics$G_means_thr[2]] <- NA
```


#### 5.3.2 PANSS overlap with other instruments 
```{r}
# items in PANSS has a mean cosine score with similar items in: 
## SANS/SAPS of: 
mean(rowMeans(m2_mat[60:89,1:59], na.rm = T)) #0.5021464
sd(rowMeans(m2_mat[60:89,1:59], na.rm = T)) #0.03358987
## DSM-5 of:
mean(rowMeans(m2_mat[60:89,90:94], na.rm = T), na.rm = T) #0.5104322
sd(rowMeans(m2_mat[60:89,90:94], na.rm = T), na.rm = T) #0.05134019
## ICD-11 of:
mean(rowMeans(m2_mat[60:89,95:101], na.rm = T)) #0.5176506
sd(rowMeans(m2_mat[60:89,95:101], na.rm = T)) #0.03932892


# items in PANSS are represented in a mean of x % of items in each instrument 
## SANS/SAPS: 
mean(rowSums(m2_mat[60:89,1:59] >0, na.rm = T))/59*100 #19.43503
sd(rowSums(m2_mat[60:89,1:59] >0, na.rm = T))/59*100 #9.259215
## DSM-5:
mean(rowSums(m2_mat[60:89,90:94] >0, na.rm = T), na.rm = T)/5*100 #43.33333
sd(rowSums(m2_mat[60:89,90:94] >0, na.rm = T), na.rm = T)/5*100 #28.32488
## ICD-11 of:
mean(rowSums(m2_mat[60:89,95:101]>0, na.rm = T))/7*100 #40
sd(rowSums(m2_mat[60:89,95:101]>0, na.rm = T))/7*100 #19.63782

# Number of items in PANSS not represented by: 
## SANS/SAPS: 
sum(rowSums(is.na(m2_mat[60:89,1:59])) == 59) #0
## DSM-5:
p_ss_not <- sum(rowSums(is.na(m2_mat[60:89,90:94]))==5) #3; P7, G4, N5
## ICD-11 of:
sum(rowSums(is.na(m2_mat[60:89,95:101]))==7) #0

```
#### 5.3.3 SANS/SAPS overlap with other instruments 
```{r}
# items in SANS/SAPS has a mean cosine score with similar items in: 
## PANSS of: 
mean(rowMeans(m2_mat[1:59,60:89], na.rm = T), na.rm = T) #0.5099867
sd(rowMeans(m2_mat[1:59,60:89], na.rm = T), na.rm = T) #0.04167022
## DSM-5 of:
mean(rowMeans(m2_mat[1:59,90:94], na.rm = T), na.rm = T) #0.5391913
sd(rowMeans(m2_mat[1:59,90:94], na.rm = T), na.rm = T) #0.07494088
## ICD-11 of:
mean(rowMeans(m2_mat[1:59,95:101], na.rm = T), na.rm = T) #0.5349388
sd(rowMeans(m2_mat[1:59,95:101], na.rm = T), na.rm = T) #0.079664


# items in SANS/SAPS are represented in a mean of x % of items in each instrument 
## PANSS: 
mean(rowSums(m2_mat[1:59,60:89] >0, na.rm = T))/30*100 #19.43503
sd(rowSums(m2_mat[1:59,60:89] >0, na.rm = T))/30*100 #13.50681
## DSM-5:
mean(rowSums(m2_mat[1:59,90:94] >0, na.rm = T), na.rm = T)/5*100 #29.83051
sd(rowSums(m2_mat[1:59,90:94] >0, na.rm = T), na.rm = T)/5*100 #24.73944
## ICD-11 of:
mean(rowSums(m2_mat[1:59,95:101]>0, na.rm = T))/7*100 #25.42373
sd(rowSums(m2_mat[1:59,95:101]>0, na.rm = T))/7*100 #21.0672

# Number of items in SANS/SAPS not represented by: 
## PANSS: 
ss_p_not <- sum(rowSums(is.na(m2_mat[1:59,60:89])) == 30) #2; SS34, SS68
## DSM-5:
ss_d_not <- sum(rowSums(is.na(m2_mat[1:59,90:94]))==5) #15; SS11, SS13, SS14, SS17, SS21, SS22, SS24, SS25, SS31, SS34, SS56, SS62, SS64, SS65, SS68
## ICD-11 of:
ss_i_not <- sum(rowSums(is.na(m2_mat[1:59,95:101]))==7) #13; SS11, SS13, SS14, SS16, SS19, SS21, SS22, SS25, SS29, SS31, SS56, SS67, SS68

```
#### 5.3.4 DSM-5 overlap with other instruments 
```{r}
# items in DSM-5 has a mean cosine score with similar items in: 
## PANSS of: 
mean(rowMeans(m2_mat[90:94,60:89], na.rm = T), na.rm = T) #0.5199992
sd(rowMeans(m2_mat[90:94,60:89], na.rm = T), na.rm = T) #0.02951063
## SANS/SAPS of:
mean(rowMeans(m2_mat[90:94,1:59], na.rm = T), na.rm = T) #0.5408937
sd(rowMeans(m2_mat[90:94,1:59], na.rm = T), na.rm = T) #0.04702941
## ICD-11 of:
mean(rowMeans(m2_mat[90:94,95:101], na.rm = T), na.rm = T) #0.6172649
sd(rowMeans(m2_mat[90:94,95:101], na.rm = T), na.rm = T) #0.02267709


# items in DSM-5 are represented in a mean of x % of items in each instrument 
## PANSS: 
mean(rowSums(m2_mat[90:94,60:89] >0, na.rm = T))/30*100 #43.33333
sd(rowSums(m2_mat[90:94,60:89] >0, na.rm = T))/30*100 #17.63834
## SANS/SAPS :
mean(rowSums(m2_mat[90:94,1:59] >0, na.rm = T), na.rm = T)/59*100 #29.83051
sd(rowSums(m2_mat[90:94,1:59] >0, na.rm = T), na.rm = T)/59*100 #7.348977
## ICD-11 of:
mean(rowSums(m2_mat[90:94,95:101]>0, na.rm = T))/7*100 #54.28571
sd(rowSums(m2_mat[90:94,95:101]>0, na.rm = T))/7*100 #15.64922

# Number of items in DSM-5 not represented by: 
## PANSS: 
sum(rowSums(is.na(m2_mat[90:94,60:89])) == 30) #0
## SANS/SAPS:
sum(rowSums(is.na(m2_mat[90:94,1:59]))==59) #0
## ICD-11 of:
sum(rowSums(is.na(m2_mat[90:94,95:101]))==7) #0

```
#### 5.3.5 ICD-11 overlap with other instruments 
```{r}
# items in ICD-11 has a mean cosine score with similar items in: 
## PANSS of: 
mean(rowMeans(m2_mat[95:101,60:89], na.rm = T), na.rm = T) #0.5198662
sd(rowMeans(m2_mat[95:101,60:89], na.rm = T), na.rm = T) #0.04432886
## SANS/SAPS of:
mean(rowMeans(m2_mat[95:101,1:59], na.rm = T), na.rm = T) #0.5422404
sd(rowMeans(m2_mat[95:101,1:59], na.rm = T), na.rm = T) #0.05301509
## DSM-5 of:
mean(rowMeans(m2_mat[95:101,90:94], na.rm = T), na.rm = T) #0.6362345
sd(rowMeans(m2_mat[95:101,90:94], na.rm = T), na.rm = T) #0.1144019


# items in ICD-11 are represented in a mean of x % of items in each instrument 
## PANSS: 
mean(rowSums(m2_mat[95:101,60:89] >0, na.rm = T))/30*100 #40
sd(rowSums(m2_mat[95:101,60:89] >0, na.rm = T))/30*100 #25.89151
## SANS/SAPS :
mean(rowSums(m2_mat[95:101,1:59] >0, na.rm = T), na.rm = T)/59*100 #25.42373
sd(rowSums(m2_mat[95:101,1:59] >0, na.rm = T), na.rm = T)/59*100 #11.57848
## DSM-5 of:
mean(rowSums(m2_mat[95:101,90:94]>0, na.rm = T))/5*100 #54.28571
sd(rowSums(m2_mat[95:101,90:94]>0, na.rm = T))/5*100 #22.25395

# Number of items in ICD-11 not represented by: 
## PANSS: 
sum(rowSums(is.na(m2_mat[95:101,60:89])) == 30) #0
## SANS/SAPS:
sum(rowSums(is.na(m2_mat[95:101,1:59]))==59) #0
## DSM-5 of:
sum(rowSums(is.na(m2_mat[95:101,90:94]))==5) #0

```



### item rep 
```{r}
#empty matrix 
m2_mat_m <- m2_mat
m2_mat[is.na(m2_mat)] <- 0
m2_mat_m[] <- 0

for (row in 1:nrow(m2_mat)){
  max_ss <- as.numeric(which.max(m2_mat[row, 1:59]))
  m2_mat_m[row, max_ss] <- m2_mat[row, max_ss] 
  
  max_p <- as.numeric(which.max(m2_mat[row, 60:89]))+59
  m2_mat_m[row, max_p] <- m2_mat[row, max_p] 
  
  max_d <- as.numeric(which.max(m2_mat[row, 90:94]))+89
  m2_mat_m[row, max_d] <- m2_mat[row, max_d] 
  
  max_i <- as.numeric(which.max(m2_mat[row, 95:101]))+94
  m2_mat_m[row, max_i] <- m2_mat[row, max_i] 
}

rowSums(m2_mat_m>0)
diag(m2_mat_m) <- NA
m2_mat_m[m2_mat_m == 0] <- NA
m2_df_m <- cor_gather(m2_mat_m)

#to incl all codes in df
ext <- merge(items[c(3:4)], nos_cat, by.x = "CODE", by.y = "ITEMS")
ext <- ext %>% slice(rep(1:n(), each = 3))

ext$rep_in <- ifelse(ext$SCALE == "ICD-11", 
       rep(c("DSM-5", "PANSS", "SANS_SAPS"), 
           each = 1), 
       ifelse(ext$SCALE == "DSM-5", 
              rep(c("ICD-11", "PANSS", "SANS_SAPS"), 
                  each = 1),
              ifelse(ext$SCALE == "PANSS", 
                     rep(c("ICD-11", "DSM-5", "SANS_SAPS"), 
                         each = 1),
                     ifelse(ext$SCALE == "SANS_SAPS", 
                            rep(c("ICD-11", "PANSS", "DSM-5"), each = 1), NA))))
colnames(ext) <- c("items", "item_name", "item_scale", "item_cat", "rep_in")
              
m2_df_m <- merge(m2_df_m, nos_cat, by.x = "var2", by.y = "ITEMS")
m2_df_m <- merge(m2_df_m, items[c(3:4)], by.x = "var2", by.y = "CODE")
colnames(m2_df_m) <- c("rep_by", "items", "cos", "rep_in", "rep_cat", "rep_name")

m2_df_m <- merge(ext, m2_df_m, by.x = c("items", "rep_in"), by.y = c("items", "rep_in"), all= T)  

m2_df_m <- merge(m2_df_m, BC_eval_df2[c(1,2,4)], by.x = c("items", "rep_by"), by.y = c("item1", "item2"), all.x = T)

m2_df_m <- m2_df_m[c(4,9,5,3, 7,6,8, 10,1,2)]
m2_df_m$cos[is.na(m2_df_m$cos)] <- 0

m2_df_m %>% 
  summarise(mean(cos), sd(cos), .by = c(item_scale, rep_in)) 
m2_df_m%>% 
  summarise(mean(cos), sd(cos), .by = item_scale)

table(m2_df_m$rep_in[m2_df_m$cos == 0], m2_df_m$items[m2_df_m$cos == 0])

(m2_df_m$items[m2_df_m$cos == 0])

table(m2_df_m$rep_in, m2_df_m$item_scale, m2_df_m$bc_overlap)

13/19*100

24/
```





